---
title: 'Linear Regression - Part II'
author: "Logan Kelly, Ph.D."
date: 'Managerial Statistics - ECON 730'
output:
  ioslides_presentation:
    highlight: tango
    widescreen: true
    theme: spacelab
  pdf_document: default
  beamer_presentation:
    slide_level: 2
    theme: metropolis
    highlight: tango
    latex_engine: xelatex
  html_document:
    css: LectureNotes.css
    highlight: tango
    number_sections: yes
    theme: lumen
  slidy_presentation:
    highlight: tango
    theme: spacelab
fontsize: 12pt
institute: University of Wisconsin-River Falls
department: College of Business and Economics
short-author: Dr. Kelly
short-date: ''
short-institute: UWRF
short-title: Statistics Lecture
classoption: aspectratio=169
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "", out.width="90%",fig.asp=.56, fig.align = "center", message=FALSE )

library(knitr)
```
<script src="https://ajax.googleapis.com/ajax/libs/jquery/3.1.1/jquery.min.js"></script>

<style>
.zoomDiv {
  opacity: 0;
  position:absolute;
  top: 50%;
  left: 50%;
  z-index: 50;
  transform: translate(-50%, -50%);
  box-shadow: 0px 0px 50px #888888;
  max-height:100%; 
  overflow: scroll;
}

.zoomImg {
  width: 100%;
}
</style>


<script type="text/javascript">
  $(document).ready(function() {
    $('slides').prepend("<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>");
    // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src'));
      $('.zoomDiv').css({opacity: '1', width: '90%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'});
    });
  });
</script>

## R Packages Used In This Lecture {-}

```{r echo=T,message=F,warning=F,results='hide'}
if (!require("tidyverse")) install.packages("tidyverse")
suppressMessages(library(tidyverse))

if (!require("Rmisc")) install.packages("Rmisc")
suppressMessages(library(Rmisc))

if (!require("psych")) install.packages("psych")
suppressMessages(library(psych))

if (!require("jtools")) install.packages("jtools")
suppressMessages(library(jtools))

```
 
 
# The Newspaper case

## Case 1: Predicting Circulation

- Case: A newspaper is considering a Sunday edition.  
  - Will they break even?  
  - How sure can we be?

- Key Concepts:                                  
  - Confidence and Prediction Intervals.
	- R-squared 

## The Decision Problem

- Financial analysis tells us that the project is good if and only if circulation is greater than 260,000.

- We need a model to predict circulation

- We have a data set consisting of circulation in 35 newspapers with various levels of circulation.

- The daily circulation of our paper is 190,000.  

## Load the data

```{r LoadCase, echo=T}
news <- read.csv("data/newspapers.csv")
describe(news, omit=T, skew=F)

print(paste("SD away from sample mean: ", (190-mean(news$Daily))/sd(news$Daily)))

```

## Scatterplot of Sunday vs. Daily Circulation

```{r echo=T, out.width="40%",fig.asp=.56}
news %>% ggplot(aes(x=Daily,y=Sunday)) +
  geom_point() +
  geom_smooth(method=lm, se=T, fullrange=F) +
  ggtitle("Sunday vs. Daily Circulation")
```

## Linear Model of Sunday Circulation - Estimation

```{r echo=T}
news.lm01 <- lm(Sunday ~ Daily, data = news)
news.lm01.resid <- resid(news.lm01)
news.lm01.fitted <- fitted(news.lm01)
```

## Linear Model of Sunday Circulation - Summary

```{r echo=T, eval=F}
library(jtools)
summ(news.lm01, digits = 3, model.info = FALSE, confint = TRUE)
```

```{r echo=F, eval=T}
library(jtools)
normal_print(summ(news.lm01, digits = 3, model.info = FALSE, confint = TRUE))
```

## Diagnostic Plots

## Residual plot from Simple Linear Model

- Actual vs. Fitted Values

```{r echo=F, fig.show="hold", out.width="35%",fig.asp=.56}
news %>% ggplot(aes(x=Sunday,y=news.lm01.fitted)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "blue") +
  xlab("Actual") +
  ylab("Fitted") +
  ggtitle("Actual vs. Fitted")
```

- Residuals vs. Daily Circulation

```{r echo=F, fig.show="hold", out.width="35%",fig.asp=.56}
news %>% ggplot(aes(x=Daily,y=news.lm01.resid)) +
  geom_point() +
  ggtitle("Residule vs. Number of Boxes")
```


## Confidence and Prediction intervals for the Dependent Variable

```{r, eval=T}
alpha <-0.05
predict(news.lm01, newdata = data.frame(Daily=190), interval="confidence", level =1-alpha)
predict(news.lm01, newdata = data.frame(Daily=190), interval="predict", level =1-alpha)

```

## Probability of circulation being greater than 190,000
- Store Prediction 
```{r, eval=T}
pred <- predict.lm(news.lm01, newdata=data.frame(Daily=190), se.fit = T, interval="predict", level =0.95)
```

## Probability of circulation being greater than 190,000
- Standard error of prediction
```{r}
crit <- qt(p= alpha/2, df = pred$df, lower.tail = FALSE)
crit
se <- (pred$fit[,"upr"] - pred$fit[,"fit"])/crit
se
```

## Probability of circulation being greater than 190,000
- Calculating the probability
```{r}
t.score <- (260 - pred$fit[,"fit"])/se 
t.score
pt(q=t.score,df=pred$df,lower.tail=F)
```


## Function to calulate forecast probability

```{r}
data(mtcars)

pred.prob <- function(lmod,new.data,alpha=0.05,alternative="greater",value){
  prediction <- predict(lmod, 
                        newdata = new.data,
                        interval="predict",
                        level =1-alpha,
                        se.fit = T)

  c.v <- qt(p=alpha/2, df = prediction$df, lower.tail = FALSE)
  
  se <- (prediction$fit[,"upr"] - prediction$fit[,"fit"])/c.v
  
  prediction$fit[,"fit"]
  t.stat <- (value-prediction$fit[,"fit"])/se
  tail <- alternative=="less"
  probability <- pt(q=t.stat,df=prediction$df,lower.tail = tail)
  return(probability)
}

pred.prob(lmod = lm(mpg ~ hp, data = mtcars),
          new.data = data.frame(hp=190),
          alpha = 0.95,
          alternative <- "greater",
          value <- 18)

```

## Conclution

There is only a 55.6% probability that the Sunday circulation will meet or exceed 260,000.


