---
title: 'Linear Regression: Dealing with Heteroscdaticity and Nonlinearity'
author: "Logan Kelly, Ph.D."
date: 'Managerial Statistics - ECON 730'
output:
  ioslides_presentation:
    highlight: tango
    widescreen: true
    theme: spacelab
  pdf_document: default
  beamer_presentation:
    slide_level: 2
    theme: metropolis
    highlight: tango
    latex_engine: xelatex
  html_document:
    css: LectureNotes.css
    highlight: tango
    number_sections: yes
    theme: lumen
  slidy_presentation:
    highlight: tango
    theme: spacelab
fontsize: 12pt
institute: University of Wisconsin-River Falls
department: College of Business and Economics
short-author: Dr. Kelly
short-date: ''
short-institute: UWRF
short-title: Statistics Lecture
classoption: aspectratio=169
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "", out.width="90%",fig.asp=.56, fig.align = "center", message=FALSE)
```
<script src="https://ajax.googleapis.com/ajax/libs/jquery/3.1.1/jquery.min.js"></script>

<style>
.zoomDiv {
  opacity: 0;
  position:absolute;
  top: 50%;
  left: 50%;
  z-index: 50;
  transform: translate(-50%, -50%);
  box-shadow: 0px 0px 50px #888888;
  max-height:100%; 
  overflow: scroll;
}

.zoomImg {
  width: 100%;
}
</style>


<script type="text/javascript">
  $(document).ready(function() {
    $('slides').prepend("<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>");
    // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src'));
      $('.zoomDiv').css({opacity: '1', width: '90%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'});
    });
  });
</script>

## R Packages Used In This Lecture {-}

```{r echo=T,message=F,warning=F,results='hide'}
if (!require("tidyverse")) install.packages("tidyverse")
  suppressMessages(library(tidyverse))
if (!require("psych")) install.packages("psych")
  suppressMessages(library(psych))
if (!require("lmtest")) install.packages("lmtest")
  suppressMessages(library(lmtest))
if (!require("sandwich")) install.packages("sandwich")
  suppressMessages(library(sandwich))
if (!require("PoEdata")) install.packages("PoEdata")
  suppressMessages(library(PoEdata))
if (!require("jtools")) install.packages("jtools")
  suppressMessages(library(jtools))
```
 
 
# Heteroscdaticity and Nonlinearity

## Recall Anscombe's Quartet
Anscombe's Quartet of ‘Identical’ Simple Linear Regressions is a set of four pairs of x/y data that all yield identical regression output. Consider the summary statistics below.

```{r}
data("anscombe")
attach(anscombe)
describe(anscombe, skew = F,ranges = F)
```

## Anscombe's Quartet Ploted

```{r echo=F, include=F}
p1 <- anscombe %>% ggplot(aes(x=x1,y=y1)) +
  geom_point() +
  theme_classic() +
  ggtitle("Plot 1") +
  geom_smooth(method="lm")

p2 <- anscombe %>% ggplot(aes(x=x2,y=y2)) +
  geom_point() +
  theme_classic() +
  ggtitle("Plot 2") +
  geom_smooth(method="lm")

p3 <- anscombe %>% ggplot(aes(x=x3,y=y3)) +
  geom_point() +
  theme_classic() +
  ggtitle("Plot 3") +
  geom_smooth(method="lm")

p4 <- anscombe %>% ggplot(aes(x=x4,y=y4)) +
  geom_point() +
  theme_classic() +
  ggtitle("Plot 4") +
  geom_smooth(method="lm")

```

<div style="float: left; width: 45%;">
```{r}
lm1 <- lm(y1~x1)
```
```{r echo=F, out.width="50%", fig.asp=0.56}
p1
```
```{r}
lm3 <- lm(y3~x3)
```
```{r echo=F, out.width="50%", fig.asp=0.56}
p3
```
</div>

<div style="float: right; width: 45%;">
```{r}
lm2 <- lm(y2~x2)
```
```{r echo=F, out.width="50%", fig.asp=0.56}
p2
```
```{r}
lm4 <- lm(y4~x4)
```
```{r echo=F, out.width="50%", fig.asp=0.56}
p4
```
</div>




## Function to Plot the Residuals vs. Predictor Plot

```{r, eval=T}
plot.resid <- function(lmod,indvar=1,title=NULL){
df = lmod$model
pr <- df %>% ggplot(aes(x=df[,indvar+1],y=resid(lmod))) +
  geom_ribbon(aes(ymin = -sigma(lmod),ymax=sigma(lmod)), fill="gray",alpha=.5) +
  geom_ribbon(aes(ymin = -sigma(lmod)*2,ymax=sigma(lmod)*2), fill="lightgray",alpha=.5) +
  geom_point() +
  theme_classic() +
  ggtitle(title) +
  ylab("Residules") +
  xlab(names(df)[indvar+1]) +
  geom_hline(yintercept=0, linetype="solid", color = "black") 
  return(pr)
}
```

## Residuals vs. Predictor Plot 

<div style="float: left; width: 45%;">
```{r echo=T, out.width="50%", fig.asp=0.56}
lm1 <- lm(y1~x1)
plot.resid(lm1,1,"Plot 1")

lm3 <- lm(y3~x3)
plot.resid(lm3,1,"Plot 3")
```
</div>

<div style="float: right; width: 45%;">
```{r echo=T, out.width="50%", fig.asp=0.56}
lm2 <- lm(y2~x2)
plot.resid(lm2,1,"Plot 2")

lm4 <- lm(y4~x4)
plot.resid(lm4,1,"Plot 4")
```
</div>

## Function to Plot a Residuals vs. Fits Plot

```{r, eval=T}
plot.fitted <- function(lmod){
df = lmod$model
title = paste("Residuals vs. Fits (Dependent Variable: ", names(df)[1], ")", sep="")
pr <- df %>% ggplot(aes(x=fitted(lmod),y=resid(lmod))) +
  geom_ribbon(aes(ymin = -sigma(lmod),ymax=sigma(lmod)), fill="gray",alpha=.5) +
  geom_ribbon(aes(ymin = -sigma(lmod)*2,ymax=sigma(lmod)*2), fill="lightgray",alpha=.5) +
  geom_point() +
  theme_classic() +
  ggtitle(title) +
  ylab("Residules") +
  xlab("Fitted") +
  geom_hline(yintercept=0, linetype="solid", color = "black") 
  return(pr)
}
```

## Residuals vs. Fits Plot

<div style="float: left; width: 45%;">
```{r echo=T, out.width="50%", fig.asp=0.56}
plot.fitted(lm1)

plot.fitted(lm3)
```
</div>

<div style="float: right; width: 45%;">
```{r echo=T, out.width="50%", fig.asp=0.56}
plot.fitted(lm2)

plot.fitted(lm4)
```
</div>

# Case 1: Ad Sales

## Ad Sales
- Increasing advertising expenditures tends to increase sales
- but at a decreasing rate.
- We can observe these diminishing returns in a scatterplot,
- and describe the relationship with a log-log model

## Ad Sales: Scatterplot
```{r, echo=F, out.width="80%", fig.asp=0.56}
ad.sales <- read.csv("data/adsales.csv")
ad.sales %>% ggplot(aes(x=exp,y=sales)) +geom_point() 
```

## Linear Model

```{r, echo=F,out.width="80%", fig.asp=0.56}
ad.sales %>% ggplot(aes(x=exp,y=sales)) + geom_point() + geom_smooth(method="lm")
```

## Linear Model:  Residuals vs. Fits Plot
```{r, echo=F, out.width="80%", fig.asp=0.56}
as.sales.lm <- lm(sales~exp, data=ad.sales)
plot.fitted(as.sales.lm)
```

## Ad Sales: Scatterplot
```{r, echo=T, out.width="60%", fig.asp=0.56}
as.sales.lm <- lm(log(sales)~log(exp), data=ad.sales)
ad.sales %>% ggplot(aes(x=log(exp),y=log(sales))) +
  geom_point() +geom_smooth(method = "lm")
```

## Linear Model:  Residuals vs. Fits Plot
```{r, echo=F, out.width="80%", fig.asp=0.56}
plot.fitted(as.sales.lm)
```

## Interpreation
- Linear Model:
  - As X increases by 1, Y increases by $\beta_X$

- Semi-Log Model:
  - As X increases by 1, Y increases by $100\times\beta_X\%$

- Log-Log Model:
  - As X increases by 1%, Y increases by $\beta_X\%$


# Case 2: Predicting Vacation Travel Distance

## Vacation Travel Distance

```{r}
data("vacation")
describe(vacation, skew=F)
```

- *miles:* miles traveled per year
- *income:* annual income ($1000)
- *age:* average age of adult members of household
- *kids:* number of children in household

## Estimate a model
```{r render = 'normal_print'}
vacation.lm <-lm(miles ~ income + age + kids, data = vacation)
summ(vacation.lm, model.info=F)
```

---
```{r, echo = F,  out.width="90%", fig.asp=0.56}
plot.fitted(vacation.lm )
```

## Residuals vs. Predictor Plot 

<div style="float: left; width: 45%;">
### income
<hr>
```{r echo=F, out.width="50%", fig.asp=0.56}
plot.resid(vacation.lm,1)
```

### age
<hr>
```{r echo=F, out.width="50%", fig.asp=0.56}
plot.resid(vacation.lm,2)
```

</div>

<div style="float: right; width: 45%;">

### kids
<hr>
```{r echo=F, out.width="50%", fig.asp=0.56}
plot.resid(vacation.lm,3)
```
</div>

## Robust Breusch-Pagan Test

- The function bptest() in package lmtest does (the robust version of) the Breusch-Pagan test in  R.

- The Null hypothesis is no heteroskedasticiy 

```{r}
bptest(vacation.lm)
```

- The small p-value implies rejection of the null hypothesis of homoskedasticity.

- This confirms our conclusion from the residual plots that heteroscadaticity is a problem.

## So now what?

- First try to address the cause of the heteroscadaticity - often non-linearity 

```{r echo=F, out.width="50%", fig.asp=0.56}
vacation %>% ggplot(aes(y = miles, x = income)) + geom_point()
```

## Report Robust Standard Errors

```{r render = 'normal_print'}
summ(vacation.lm, model.info=F, robust=T)
```

## Comparing Output

```{r results="hide"}
OLS <- summ(vacation.lm, model.info=F, robust=F)
ROBUST <- summ(vacation.lm, model.info=F, robust=T)
export_summs(OLS,ROBUST, model.names = c("OLS SE","Robust SE"))
```

##

```{r echo=F}
export_summs(OLS,ROBUST, model.names = c("OLS SE","Robust SE"))
```







