---
title: "Tools for summarizing and visualizing regression models"
author: "Jacob Long"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: "spacelab"
vignette: >
  %\VignetteIndexEntry{Tools for summarizing and visualizing regression models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo=FALSE}
required <- c("survey", "huxtable", "broom", "lme4", "quantreg")
if (!all(sapply(required, requireNamespace, quietly = TRUE)))
  knitr::opts_chunk$set(eval = FALSE)
knitr::opts_chunk$set(message = F, warning = F, fig.width = 6, fig.height = 4,
                      dpi = 100, render = knitr::normal_print)
library(jtools)
```

The support `jtools` provides for helping to understand and report the results
of regression models falls into a few broad categories:

* Generating flexible table output in the console that includes multiple
standard error specifications, toggles for confidence intervals, VIFs, p values,
and so on (`summ`)
* Plotting predicted data from models to aid in substantive interpretation
and understanding model fit, including models with interactions (`effect_plot`;
see other vignette)
* Plotting regression coefficients and their uncertainty in a visually appealing
way (`plot_coefs`, `plot_summs`)
* Exporting regression summaries as tables in PDF/LaTeX and Word formats for
publication (`export_summs`)

# `summ`

When sharing analyses with colleagues unfamiliar with R, I found that the output
generally was not clear to them. Things were even worse if I wanted to give
them information that is not included in the `summary` like robust
standard errors, scaled coefficients, and VIFs since the functions for 
estimating these don't append them to a typical regression table. After creating
output tables "by hand" on multiple occasions, I thought it best to pack things
into a reusable function: It became `summ`.

With no user-specified arguments except a fitted model, the output of `summ`
looks like this:

```{r}
# Fit model
states <- as.data.frame(state.x77)
fit <- lm(Income ~ Frost + Illiteracy + Murder, data = states)
summ(fit)
```

